|

### 2021-08-04 article/article.Rmd
changes article and appendix - adding model 1

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/legislature-size-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find no robust evidence for the effect of legislature size on government budgets. Yet the aggregate results show significant heterogeneity. While earlier studies provide moderate support for the \"law of $1/n$\", papers using causal inference methods consistently find a negative relationship between seats and spending. The available evidence also indicates that proportional representation and mixed voting systems are no more likely to overspend than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# Clean up
rm(list = ls())

## Starting
set.seed(732578) # From random.org

# Needed packages
pkgs <- c("tidyverse", "meta", "metafor",
          "readxl", "data.table",
          "knitr", "gridGraphics", "gridExtra",
          "ggpubr", "kableExtra", "magick")

# Install if not already installed
installIfNot <- function(x) {
  if (x %in% rownames(installed.packages()) == FALSE)
    install.packages(x, dependencies = T,
					 repos = "http://cran.us.r-project.org")
}
lapply(pkgs, installIfNot)

# Load packages
lapply(pkgs, require, character.only = T)
devtools::install_github("isubirana/compareGroups")
library("compareGroups")

# Load datasets
load("../dataset/dataCoefs.RData")

# Build plot function for forest plots
build_forest <- function(mod, capt, lsize = 22, ttl = NULL) {
  # Build dataset for plot
  mod2 <- tibble(
    TE = mod$TE,
    seTE = mod$seTE,
    studlab = mod$studlab,
    lower = mod$lower,
    upper = mod$upper,
    group = "A") %>%
    bind_rows(.,
              aux = tibble(
                TE = c(mod$TE.random, NA),
                seTE = c(mod$seTE.random, NA),
                studlab = c("Overall Effect",
                            "Prediction Interval"),
                lower = c(mod$lower.random,
                          mod$lower.predict),
                upper = c(mod$upper.random,
                          mod$upper.predict),
                group = "B")) %>%
    group_by(studlab) %>%
    mutate(studlab2 = paste0(studlab, "_", 1:n())) %>%
    ungroup()

  # Graph limits
  limg <- max(abs(c(mod2$lower, mod2$upper)))

  # Build plot
  p <- mod2 %>%
    ggplot(aes(y = reorder(studlab2, TE),
               x = TE, xmin = lower, xmax = upper)) +
    geom_point(aes(color = group)) +
    geom_errorbarh(aes(color = group),
                   height = 0.1) +
    scale_color_manual(values = c("#000000", "#8b0000")) +
    scale_x_continuous(limits = c(-1.1 * limg, 1.1 * limg)) +
    scale_y_discrete(
      labels = function(x)
        str_replace(x, "_[0-9]*$", "")) +
    geom_vline(xintercept = 0,
               color = "#000000", linetype = "dashed") +
    labs(x = "",
         y = "") +
    facet_grid(group~., scales = "free", space = "free") +
    labs(caption = capt,
         title = ttl) +
    theme_minimal() %+replace%
    theme(strip.text.y = element_blank(),
          legend.position = "none",
          axis.text.y = element_text(size = .8 * lsize,
                                     hjust = 1),
          axis.text.x = element_text(size = .6 * lsize,
                                     hjust = 1.1),
          plot.caption = element_text(size = lsize),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5,
                                    face = "bold",
                                    margin = margin(0, 0, 10, 0)),
          panel.grid.major = element_blank())
  return(p)
}

# Build forest plot for heterogeneous analysis
# Build plot function for forest plots
build_forest_het <- function(mod, capt, lsize = 22, ttl = NULL, hetvar = NULL) {
  mod <- update(mod, byvar = hetvar, print.byvar = F)

  # Build dataset for plot
  mod2 <- tibble(
    byvar = mod$byvar,
    TE = mod$TE,
    seTE = mod$seTE,
    studlab = mod$studlab,
    lower = mod$lower,
    upper = mod$upper,
    group = "A") %>%
    arrange(byvar)
  auxmod <- tibble()
  for (i in rev(mod$bylevs)){
    auxmod <- rbind(auxmod, tibble(byvar = i,
                    TE = NA,
                    seTE = NA,
                    studlab = toupper(i),
                    lower = NA,
                    upper = NA,
                    group = "B"))
    auxmod <- rbind(auxmod, 
                    filter(mod2, byvar==i) %>% 
                      arrange(desc(TE)))
    auxmod <- rbind(auxmod, tibble(
      byvar = i,
      TE = mod$TE.random.w[which(mod$bylevs==i)],
      seTE = mod$seTE.random.w[which(mod$bylevs==i)],
      studlab = 'Subgroup Effect',
      lower = mod$lower.random.w[which(mod$bylevs==i)],
      upper = mod$upper.random.w[which(mod$bylevs==i)],
      group = "B"))
  }
  auxmod <- rbind(auxmod, tibble(
    byvar = NA,
    TE = c(mod$TE.random, NA),
    seTE = c(mod$seTE.random, NA),
    studlab = c("Overall Effect", "Prediction Interval"),
    lower = c(mod$lower.random, mod$lower.predict),
    upper = c(mod$upper.random, mod$upper.predict),
    group = "B"))
  mod2 <- data.frame(auxmod)
  mod2$byvar <- toupper(mod2$byvar)
  TEaux <- mod2$TE
  TEaux[mod2$studlab== 'Subgroup Effect'] = TEaux[mod2$studlab== 'Subgroup Effect'] - 100

  # Graph limits
  limg <- max(abs(c(mod2$lower, mod2$upper)))

  # Build plot
  p <- mod2 %>%
    ggplot(aes(y = reorder(studlab, TEaux),
               x = TE, xmin = lower, xmax = upper)) +
    geom_point(aes(color = group)) +
    geom_errorbarh(aes(color = group),
                   height = 0.1) +
    scale_color_manual(values = c("#000000", "#8b0000")) +
    scale_x_continuous(limits = c(-1.1 * limg, 1.1 * limg)) +
    scale_y_discrete(
      labels = function(x)
        str_replace(x, "_[0-9]*$", "")) +
    geom_vline(xintercept = 0,
               color = "#000000", linetype = "dashed") +
    labs(x = "",
         y = "") +
    facet_grid(byvar~., scales = "free", space = "free") +
    labs(caption = capt,
         title = ttl) +
    theme_minimal() %+replace%
    theme(strip.text.y = element_blank(),
          legend.position = "none",
          axis.text.y = element_text(size = .8 * lsize,
                                     hjust = 1),
          axis.text.x = element_text(size = .6 * lsize,
                                     hjust = 1.1),
          plot.caption = element_text(size = lsize),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5,
                                    face = "bold",
                                    margin = margin(0, 0, 10, 0)),
          panel.grid.major = element_blank())
  return(p)
}
```

[^huzeyfe]: Research Associate, Contemporary Brazilian History Research and Documentation Center, School of Social Sciences, Fundação Getulio Vargas, <huzeyfealptekin@gmail.com>.

@@

[^catarina]: PhD Student, Department of Political Science, University of California San Diego, <catarinamroman@gmail.com>, <http://catarinaroman.github.io>.

@@

Over the past decades, a large literature has examined the relationship between
legislature size and public expenditure. @weingast1981political provided the
general framework to analyse distributive politics. The authors argue that the
larger the number of legislative districts ($n$), the smaller the share of tax
burden each one will bear ($1/n$), thus legislators have an incentive to
overspend in their districts and transfer the costs to the entire polity. Early
studies that empirically tested the "law of $1/n$", as the theory is currently
known, indeed found a positive correlation between the number of legislature
seats and different measures of government spending, although these first
results were mainly based on US state legislatures and the effect was often
limited to one house [e.g., @baqir2002districting; @gilligan1995deviations;
@gilligan2001fiscal].

Later research, however, has questioned the validity of the "law of $1/n$".
@primo2008distributive affirm that, due to spatial spillovers, a collection of
small districts can supply public goods more efficiently than the central
government. The authors conclude that a "reverse law of $1/n$" may hold,
wherein a higher number of legislators in small constituencies decrease the
overall public spending. Similarly, @primo2006stop and @chen2007law find that
lower and upper chambers may have mixed effects on government spending, while
@petterssonlidbom2012size argues that the impact of larger chamber sizes is
negative when using data from Finland and Sweden.

Since many empirical tests of the "law of $1/n$" have produced conflicting
results, scholars have expanded this research agenda and closely investigated
how institutional factors condition the original formulation of the theory. For
instance, authors such as @crowley2019law and @pecorino2018supermajority
accurately point out that collective action problems have been overlooked in
the literature, and recent findings indicate that bicameralism
[@maldonado2013legislatures], intergovernmental competition
[@crowley2015local], redistricting [@lee2018court], and party ideology
[@bjedov2014impact] strongly influence the relationship between seats and
spending. Moreover, the literature has increasingly applied causal inference
designs to estimate the effect of the "law of $1/n$", and in contrast to
previous studies using panel data, regression discontinuity designs generally
indicate that more legislators decrease public expenditures
[@debenedetto2018effect; @hohmann2017effect; @lewis2019legislature]. In this
respect, scholars have long been aware of the theoretical and empirical
limitations of the "law of $1/n$", and the proliferation of new studies reflect
a conscious attempt to assess the robustness of the theory [@crowley2019law].

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to evaluate the impact of legislature size over
government spending across several dimensions. Our study sample mirrors the
diversity of the literature. Out of the `r dim(dat)[1]` coefficients
included in our main analysis, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Significant')])`\% of
them are positive and statistically significant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Insignificant')])`\%
are positive and statistically insignificant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Insignificant')])`\%
are negative and statistically insignificant, and `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Significant')])`\% are
negative and statistically significant. Given the volume and the disparity of
the studies, we employ meta-analysis to summarise the results. Meta-analysis
provides a rigorous approach to combine heterogeneous outcomes into a single
estimation, and it allows scholars to gain valuable insights from the
aggregated data [@cooper2019handbook; @hedges1985statistical]. Meta-analysis
can also identify potential sources of study variability, enabling researchers
to assess threats to external validity and direct future efforts into more
promising areas of academic inquiry [@doucouliagos2008democracy]. Research
synthesis methods have been successfully applied in medicine and psychology
since the 1970s [@glass2015meta], and our work contributes to the burgeoning
literature that uses meta-analytic methods to understand challenging questions
in political science [@costa2017responsive; @doucouliagos2008democracy;
@green2013field; @lau2007effects; @schwarz2020supporting].

We run binomial tests and meta-regressions to further evaluate the relationship
between legislature size and government expenditures. We use binomial tests
to assess whether the proportions of positive and negative coefficients in our
sample are statistically different from each other. In the meta-regressions, we
measure the effect of five moderators that capture different sources of
heterogeneity in the literature. As our sample includes both published and
unpublished articles, we also construct a series of funnel plots to test for
publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs, we see that all three
papers included in our sample suggest that a higher number of legislators leads
to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature]. In this regard, it is possible that methodological
choices partially explain the divergent results observed in the literature, as
studies that use causal inference methods consistently point to the same
direction. One limitation of this finding is that these studies only test the
impact of lower house size on the natural logarithm of expenditure per capita,
thus it remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using our extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Since most unicameral
legislatures are located in local governments (municipalities or districts),
this result supports the theory strictly in its first formulation. Moreover,
our meta-regressions indicate that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests. Non-majoritarian voting systems decrease government spending as
a percentage of GDP, but they do not have a relevant impact on the other
measures of public expenditure, following the idea that the $1/n$ effect grows
weaker as the empirical cases distance from the original definition of the law.
Finally, meta-regression results confirm that regression discontinuity models
reduce public spending estimates.

In summary, our results are in line with previous findings that lend partial
support to the law of $1/n$ [@chen2007law; @crowley2015local; @primo2006stop].
An assessment of the effects of collective action costs in these analyses is
crucial to determine the full mechanism operating behind these results.
Overall, our analysis does not provide robust support for the theory as
originally conceived.

# Data and Methods

We compiled our study sample in three search rounds. In the first round, we
gathered data from three large academic databases (Scopus, Microsoft Academic,
and Google Scholar) and looked for studies that were published in English and
cited \citet{weingast1981political}, as it is the foundational work in the
literature on the "law of $1/n$". To ensure that our sample was comparable, we
only selected papers that used quantitative methods to analyse
data\footnote{Since meta-analysis requires a single estimate per observation,
we excluded articles that use interaction terms or quadratic specifications of
our selected variables. Please refer to Section C in the Supplementary Material
for a detailed description of the selection procedure. We also included two
PRISMA (Preferred Reporting Items for Systematic Reviews and Meta-Analyses)
flow diagrammes \citep{liberati2009prisma} showing the number of resulting
papers after each review step.}. After this stage, we identified six
measurements that the literature often employs to quantify government
expenditure and legislature size. For government expenditure, our study sample
uses (i) public expenditure as a share of GDP; (ii) public expenditure per
capita; and (iii) the natural logarithm of public expenditure per capita as its
main variables of interest. In regards to legislature size, the variables are
(i) lower chamber size; (ii) natural logarithm of lower chamber size; and (iii)
upper house size\footnote{There are a few important nuances concerning coding
of these variables. Unicameralism, for example, is captured both by lower
chamber size ($n = 7$) and by log lower chamber size ($n = 5$). Since much of
the literature estimates how institutional designs affect this relationship,
ours and many other articles use both lower and upper chamber sizes as main
explanatory variables. We did not find any article that used the natural
logarithm of upper chamber size in their models.}.

In the second round, we did not require articles to cite @weingast1981political
and used a keyword-based query on Google Scholar to broaden the scope of the
first search. The search string contained terms strongly associated with the
literature on the "law of $1/n$" and was as follows: `("upper chamber size" OR
"lower chamber size" OR "council size" OR "parliament size" OR "legislature
size" OR "number of legislators" OR "legislative size") AND ("spending" OR
"expenditure" OR "government size")`. We again restricted the search to
articles written in English which employed quantitative methods. This search
added two new results to our sample [@coate2011government;
@debenedetto2018effect], but neither of them included variables beyond the six
measures we had previously identified. In the third search round, we looked
into the personal webpages of every author whom we had already included in our
sample. The purpose of this manual search was to assess whether there was any
working paper or unpublished manuscript that we had missed in the two former
queries. All papers we found in the last search had already been included in
our sample. Combined, the three searches produced a dataset of 
`r length(unique(dat$id))` studies as of the 10\textsuperscript{th} of March 2021.
Table \ref{tab:papers} contains the full list of articles we analyse in this
paper.

>>>>

---
title:  |
        | Legislature Size and Welfare:
        | Evidence from Brazil
author: 
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
date: \today
thanks: "We thank Cedric Antunes, Victor Araujo, Eric Arias, Taylor Boas, Natalia Bueno, Andrea Freitas, Saad Gulzar, James Mahon, Lucas Novaes, Nelson Ruiz, Catarina Roman, David Skarbek, and Denis Stukal for their valuable comments. We also thank the participants of seminars and conferences at CEPESP, CUNY, INSPER, IPSA, LACEA, LASA, MPSA, Notre Dame, NYU, PUC Chile, SPSA, and Uppsala. We gratefully acknowledge the financial support of the São Paulo Research Foundation (FAPESP), grant number 2017/07290-5. All materials required to replicate this study are available at <https://github.com/umbertomig/legislature-size-welfare>."
abstract: "What is the effect of legislature size on social welfare? While the literature suggests that increasing the number of legislators broadens political representation and raises government expenditures, its implications for public service provision remain understudied. In this paper, we exploit exogenous changes in city-council size in Brazil and show that adding a legislator increases education and healthcare indicators. XXXX Education quality and preventive health care remain unaffected, but primary school enrolment and infant mortality significantly improve. To investigate the mechanism, we surveyed former councillors and analysed 346,553 bills proposed between 2005 and 2008. This analysis shows that politicians prefer to provide private and targeted goods instead of public goods. This paper has implications for the design of legislative institutions."
abstractspacing: double
keywords: Legislature size; Legislative politics; Public goods provision; Regression discontinuity design; Brazil
jelcodes: HXXXX
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide'}
# XXXX
```

@@

[^gabriel]: Associate Professor, Department of International Relations, São Paulo State University, <gabi.cepal@gmail.com>.

@@

Legislative institutions are crucial for welfare and service provision in democratic countries. Historically, the division of power and the birth of a representative government placed the legislatures at the center of popular consent \citep{north1989constitutions}. In modern democracies, most countries have assemblies at all levels of government. These institutions affect the information and quality of legislation \citep{krehbiel2010information}, check and veto the other political powers \citep{tsebelis2002veto}, and influence policymaking \citep{weingast1988industrial, north1989constitutions, auriol2012optimal}. Moreover, they also determine taxation, government expenditure \citep{Weingastetal1981, PrimoSnyder2008, stasavage2011states}, and oversee service provision \citep{mccubbins1984congressional, poulsen2018political}.

Regardless of differences in electoral rules, social composition, and procedures, all legislatures share one feature: more than one representative composes them. Legislature size relates to collective action problems \citep{crain1979cost, rogers2002free}, government spending \citep{Weingastetal1981, PrimoSnyder2008}, and representation \citep{allenrepresentation}, but how these components affect welfare remain understudied. On the one hand, enlarged legislatures increase government spending and lawmaking, and this may improve some services \citep{Weingastetal1981, }. On the other hand, adding a legislator can increase the number of veto players and broaden the collective action problems, hindering service provision. Therefore, larger legislatures have an ambiguous effect on welfare, and XXXX.

In this paper, we propose a theory of legislature size and welfare that takes into account two crucial components: the XXXX and the partisan affiliation of the legislators.

XXXX Umberto: Parei aqui.

This paper fills this gap by studying the effects of legislature size on service provision. Assuming that politicians want to further their careers, we argue that two variables determine service provision: the perceived importance of the service by voters and the politician's capacity to claim credit for its provision. Politicians perform a simple cost-benefit analysis: if providing a service results in high electoral yield, and it is easy to claim credit for, politicians will produce this service abundantly. When a service is hard to claim credit for, politicians only invest their time if it is highly preferred by voters. Finally, when services are both hard to claim credit for and of low importance to voters, they are placed at the bottom of the politicians' priorities.

To test this theory, we exploit an exogenous variation in Brazilian city council sizes. In 2004, the judiciary reinterpreted the constitution, unexpectedly changing the number of city councilors in Brazilian municipalities. Before the decision, towns could freely choose their legislature size, but in March 2004, the Superior Electoral Court created population thresholds, adding one councilor for each set of 47,619 inhabitants. Around the cutoffs, this decision represents an exogenous change in city council size, allowing us to investigate the welfare effects of increasing the legislatures in Brazil.

Studying the 2005 to 2008 Brazilian mayoral term, when the ruling was in effect, we find improvements in health care and education. Increasing the council size by one legislator lowers infant mortality by 1.99 per one thousand infants born and increases elementary school enrollment by 2.58 children in kindergarten classrooms. However, the additional legislator had no influence on preventive health care and education quality. [] Voters place a high electoral premium on hospital care and mediating school enrollment is relatively easy to claim credit for. In line with our theoretical model, these services should improve with legislative size. On the other hand, preventive health care is harder to claim credit for, and school quality improvements are expensive and pay off only in the long-run. Thus, these services remain unchanged with larger legislatures.

Finally, we present three sets of evidence that the results are consistent with the micro-level representation practices of city councilors. First, we show that the new councilor has access to government resources and uses public office appointments to mediate service provision. They have a 91\% chance of belonging to the mayoral coalition, and they appoint an average of 105 extra bureaucrats for many positions at the mayor's office, having loyal workers inside the municipal bureaucracy. Second, we analyzed 346,553 bills approved by city councilors in municipalities close to the council size thresholds. In line with our theory, we find that councilors prefer to provide targeted goods (72.9\% of all bills), as they satisfy the populations' needs and are easier to credit claim. Finally, we ran an online survey among the councilors during the period, showing that most councilors believe that voters prefer targeted services (68.8\%) to lawmaking (24.6\%) and oversight (30\%).\footnote{Councilors were able to select multiple categories, so the sum can be over 100\%.} We also refute three alternative explanations: increased representation of women and non-white legislators, change in electoral competitiveness, and a larger average of approved legislation by each councilor.

# Legislature Size, Legislative Support, and Welfare
\label{sec:theory}

What is the effect of larger legislatures on public service provision? The most striking feature of a legislature is the fact that decisions are undertaken collectively. Legislators have to propose changes in the status-quo, agree upon what they want to decide, evaluate proposals put forward by mayors, discuss and improve the proposals, and decide which changes they will carry on.

In this sense, legislature size is critical in determining the efficiency of the decision-making process. Upon adding a new city councilor, the most simple change is the increased production capability. Legislative production, analogous to what happens in firms, may increase when adding an extra legislator \citep{crain1979cost, crain1982team, rogers2002free}. For instance, suppose that working full time, politicians can approve five bills a month. Ten politicians can together approve 50 bills, while 11 can pass 55. This increases the number of measures adopted, and if laws enact services to constituencies, then we shall witness an increase in service provision.\footnote{An alternative explanation relates to the Condorcet Jury Theorem. In these models, there is a binary choice, with one of them being ex-ante the best. If legislators access the best choice with a probability higher than 0.5, more legislators will increase the change of selecting the best choice \citep{myerson1998extended}. However, this logic assumes that politicians choose among exogenously provided proposals, which makes sense in the case of a jury, but is somewhat uncommon for legislatures. \citet{auriol2012optimal} compute the optimal size of chambers considering that this is the solution to the interplay between specialization costs and citizens' preferences.}

However, increasing the size of legislatures also carries considerable costs. Larger legislatures can raise the transaction costs \citep{weingast1988industrial}, increase collective action problems \citep{crain1979cost}, and make  forming majorities more costly \citep{crain1982team}. These negative features, together with the fact that legislators may free-ride on each others' proposals, may decrease legislative productivity. For instance, if decisions need a majority or qualified majority to be approved, more legislators create a perverse incentive, increasing the difficulty of approval. Moreover, legislators are pressured to provide targeted services for their constituencies at the expense of the entire polity \citep{Weingastetal1981, PrimoSnyder2008}, and this can raise inefficiencies in expenditure allocation. Additionally, transaction costs within legislatures may increase the difficulty of passing bills \citep{weingast1988industrial, baron1989bargaining}. Therefore, it is unclear whether and how larger legislatures impact lawmaking and service provision.

In our view, to understand the effects of larger legislatures, we need to consider the political motivations behind the legislator's work. While an extra politician may increase the productivity frontier of a legislature, the types of services that will be favored remain unclear. To understand the political motivations, we need a theory that accounts for the cost-benefit calculations performed by politicians.

When deciding which service to prioritize, politicians look into two dimensions of the service provision. First, the competition with other legislators motivates politicians to provide services that they believe voters want, regardless of whether these policies improve long-term welfare. For instance, suppose that there are three services, organized hypothetically based on beliefs about which of these three voters are more likely to reward:\footnote{Note that there is a difference between \textit{what voters want} and what \textit{politicians believe voters want}. Although politicians and brokers mostly access voter's beliefs accurately \citep{FinanSchechter2012, Stokesetal2013}, the crucial signal that a politician gets is vote shares, which are aggregated. It is hard to disentangle which service generated the highest electoral yield for the politician. Therefore, the politician's beliefs play an essential role in the services that they prioritize.} improving a local health clinic (most preferred), building a playground (second most preferred), or increasing the quality of education (least preferable). Politicians seeking recognition will mostly want to improve the health care system, which they believe will most benefit their electoral yields. After getting this done, they will shift efforts toward the playground construction and finally to education quality. Therefore, increasing legislature size will disproportionately increase the provision of services that the politicians believe voters prefer and will reward the most in the polls.

Second, to get reelected, besides providing the services that voters prefer, politicians have to take credit for facilitating provision \citep{fiorina1987personal, gulzar2017politicians, nielsen2017politicians, silva2017clarity}. Typically, the services that are easiest in terms of credit claiming are either providing local public goods, such as improving a neighborhood health clinic, or purely clientelistic and personalistic services \citep{Weingastetal1981, kuschnir2000cotidiano, Stokes2005, nichter2011electoral, Stokesetal2013, luna2014segmented, vieira2015clientelismo, bertholini2018pork}. This explains why clientelism and pork-barrel are persistent in democratic polities. Table \ref{tabcli} summarizes the incentives.

\begin{table}[htp!]
\centering
\caption{Effects of Legislature Size on Service Provision}
\label{tabcli}
\scalebox{1}{%
\begin{tabular}{@{}lcc@{}}
\toprule
& \textbf{Hard for Credit Claim} & \textbf{Easy for Credit Claim} \\ 
\midrule
\multirow{3}{*}{\textbf{Low ranked by voters}} & No changes & None or mild \\
& or deterioration & improvements \\
& (Preventive Health)  & (Honors bills) \\[0.5em]
\multirow{3}{*}{\textbf{High ranked by Voters}} & Mild to low & Mild to strong \\ 
& improvement & improvement \\
& (Infant Mortality)  & (School Enrollment) \\
\bottomrule
\end{tabular}
}
\end{table}

These two dimensions provide a simple yet powerful guide of the politicians' incentives. When voter places a high premium on the provision of a given service, and it is easy to claim credit for the provision, politicians will concentrate their efforts in providing it. The more politicians, the higher the provision of the service. Consider, for instance, constituency service, such as getting a child enrolled at a public school. The voters highly rank service such as this, and the politicians can break the school access directly, getting all the credit for the provision. Therefore, these services tend to increase, up to the point of being over provided.

When a service is hard to credit claim, to be provided, it has to be highly ranked by the voters. Consider services that decrease infant mortality. Although policies in this direction might be harder to be directly associated with one given politician, every politician wants to be known as the ones that diminished infant mortality. The death of a child has devastating effects over a polity, and no politician wants to be associated with it. On the other hand, every parent would be extremely thankful for the politicians that improved health care and saved their child. Moreover, politicians might facilitate access to medication and hospitals, which in turn is easier for credit claiming. Therefore, these services will have anywhere from modest to substantial improvements.

Suppose now that service is secure to credit claim for, but gives low electoral yield. In this case, although the provision can be unquestionably traced back to the legislator, voters care little about the provision. For instance, honors and appraisal legislation are very common in Brazilian municipalities. However, they are perceived by politicians as low electoral yield, but the legislator that proposed the homage usually gets associated with it. These types of legislation increase with legislature size, but their effect on welfare is negligible.

Finally, services that are difficult to credit-claim, or that politicians believe that voters care very little about, are expected to receive little invested effort from politicians. For instance, consider preventive health care. Preventive care refers to health measures, such as prenatal care or vaccination, undertaken to lower the chances or the severity of a disease(s) in the future. However, it is a compounded chance: the voter has to consume the service, but its effectiveness depends on voters getting sick. Nowadays, we are witnessing how hard it is to vaccinate children, despite all the scientific evidence demonstrating its effectiveness. Moreover, rational voters may fail to punish politicians for low investments in preventive public policies, such as natural disasters or climate change mitigation \citep{gailmard2019preventing}. Hence, voters might also not punish politicians for low investments in preventive health care because they might only observe whether a politician properly invested in preventive public policies in cases of emergency. Therefore, we shall expect little to no improvement in these types of services with larger legislatures.

In sum, these considerations suggest that instead of looking for overall positive or negative effects, the effects of legislature size should manifest diversely, conditional on the type of service studied. The most significant changes should be expected in services that voters want and that are easier to claim credit for. Services with only one of these characteristics will be intermediate in terms of their priority and efforts exerted by politicians in their provision and will improve with less intensity. Services hard to credit claim and low ranked by voters will tend to be unaffected or under-provided by lawmakers.

# Background on the Brazilian Case
\label{sec:background}

In 2004, Brazil comprised 5,560 municipalities. According to the Brazilian Constitution, each city must provide healthcare, primary education, transportation, and infrastructure to its citizens. Municipalities have the authority to enact local laws and collect taxes on housing and services. However, their room for fiscal manipulation is limited. On average, municipal taxes account for only 5\% of the local budgets, while mandatory transfers from state and federal governments correspond to the remaining 95\%. This reduces the ability of municipal legislators to effectively influence local spending, what makes improvements in local services dependent only on the efforts of city-level politicians. XXXX

With regards to their political organisation, municipalities have one mayor and one city council. Citizens choose their representatives by direct vote in the same election. Mayors are in charge of organising the provision of public services, as well as proposing laws and amendments to the tax code. Proposals submitted by the mayor are subject to the approval of the city council. Councillors, in turn, have two primary duties. First, they should discuss and vote legislation put forward by the mayors or by their peers. Second, they oversee the public budget and sanction mayors who do not comply with fiscal legislation. Councillors also provide an array of public and personal services for their constituents, such as helping voters access public hospitals, facilitating school enrolment, and even paying their voters' utility bills \citep{nichter2011electoral, AlmeidaFelix2012}.

Until 2004, there was no specific regulation on the size of city councils. The 1988 Brazilian Constitution set only broad guidelines about the number of council seats in each municipality, which allowed cities to determine how many legislators they should have in their local chambers. This led to severe imbalances in the proportion of representatives per citizen across the country. For instance, the Nova Russas municipality had 30,009 inhabitants and 21 councillors, which accounts for one legislator per 1,429 citizens. In contrast, the city of Sorocaba had only 15 councillors for 528,735 inhabitants, that is, one legislator per 35,249 citizens. In 2003, this malapportionment problem gained national visibility when Mira Estrela, a small municipality with only 2,651 residents, reduced its city council from 11 to 9 seats because of fiscal considerations. The change motivated a legal dispute that reached the Brazilian Supreme Court. The Court favoured the city's decision and ruled that the size of every local council should correspond to precise population thresholds.

Following the ruling of the Supreme Court, in March 2004 the Brazilian Electoral Court established a series of population cutoffs for all municipal legislatures. The Court specified that, following the Brazilian Constitution, cities with up to one million residents should have between 9 and 21 legislators. The number of seats was allocated as follows. Each city starts with 9 councillors, then adds one legislator for every 47,619 inhabitants until their councils reach 21 members. This threshold implies that cities with a population between 571,428 and one million should all the same number of legislators, that is, 21 of them. Conversely, cities between one and four million inhabitants should have at least 33 councillors, and this number increases up to 41 members following the previous formula of one additional legislator per 47,619 inhabitants.

This decision makes Brazil the ideal testing ground for our theory. Although an endogenous decision motivated the Electoral Court ruling, the population cutoffs created sharp discontinuities in the distribution of local council seats. That is, cities close to the cutoffs had very similar characteristics, but those which were just above the threshold gained a new legislator. Moreover, cities could not self-select into any group, as their population estimates were calculated using the 2003 projections by the Brazilian Census Bureau (IBGE). This change in council size also happened only seven months before the election, making it impossible for candidates to adjust their service provision strategies for the 2004 election. Therefore, the ruling allows us to study the effect of larger legislatures on public service provision during the 2005-2008 term while holding all else constant.

\section{Empirical Strategy}
\label{sec:empstrat}

\subsection{Methodology}
\label{sub:methods}

We employ a regression discontinuity design (RDD) using the population thresholds established by the 2004 Superior Electoral Court resolution as treatment indicators. Figure \ref{fig:runningvarplot} displays the distribution of municipalities within each city council size. These population thresholds remained in effect until 2009, when Congress amended the Brazilian Constitution. This study covers the period between 2005 and 2008, which constitutes the full mayoral and city-council term following the 2004 Brazilian elections.
